---
title: "Assignment4"
author: 'Team 2: Riti Dabas, Anoushka Mahar, Dylan Koury'
date: "12/2/2021"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Part 1
```{r, echo=F}
PStar = 1.5 #Long run equilibrium price
Phi = -0.4 #Persistence factor
n = 10000 #number of prices
meanEr = 0 #Mean error
stdDvEr = 0.08 #Std Deviation error
error = rnorm(n=n, mean = meanEr, sd=stdDvEr)
Prices = c() #List of prices
P0 = PStar #Setting initial price to equilibrium price
for (i in 1:n) {
  pt = PStar*(1-Phi) + Phi*P0 + error[i]
  Prices = c(Prices, pt)
  P0 = pt
}
```

##Part 2
```{r, echo=FALSE}
plot(ts(Prices[1:50]), main="First 50 Prices", ylab="Price", xlab="Time (unspecified)")
abline(h=mean(Prices[1:50]), col="dark green")
```
Looking at the graph, the prices appear stationary as there is an even distributon around the mean across all 50 temporal points.

DISCUSS HOW PROCESS WOULD CHANGE WITH PHI 0 OR 1

HOW EFFECTIVE CAN WE BE AT PREDICTING PRICES IN THESE CASES

##Part 3
```{r, echo=FALSE}
library(forecast)
Acf(Prices,main="MA(1), theta=0.05", lag.max=10)
Pacf(Prices,main="MA(1), theta=0.05", lag.max=10)
```
The ACF shows us that there are at least 4 statistically significant correlations with a couple borderline correlations. These 

#Part 4
```{r, echo=FALSE}
library(dynlm)
options(scipen=999)
PricesTS <- dynlm(Prices~L(Prices,-1))
summary(PricesTS)
```
INTERCEPT: -0.00000000000000255795 
SLOPE: 1.00000000000000177636 

##Part 5
```{r, echo=FALSE}
library(tidyverse)
avocado <- read_csv("avocado.csv")
Avocado_Prices <- ts(avocado$Avocado_Prices,frequency=52,start=c(2018,1))
```

##Part 6
```{r, echo=FALSE}
plot(Avocado_Prices, xlab="Year", ylab="Price ($)", main="Avocado Prices (2018-20)")
abline(h=mean(Avocado_Prices),col="dark green")

```







